\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{array}
\usepackage{fancyhdr}
\usepackage{amssymb}
\usepackage[shortlabels]{enumitem}

\pagestyle{fancy}
\fancyhead[L]{Banghao Chi}
\fancyhead[C]{Homework 3}
\fancyhead[R]{15th Feb}

\fancyfoot[C]{\thepage}

\renewcommand{\headrulewidth}{0.5pt}
\renewcommand{\footrulewidth}{0.5pt}

\begin{document}

\section*{Exercise 1}
Calculate the order of each element in $\mathbb{Z}_{18}^*$. \\

\textbf{Solution:} \\

From Homework 2, we know that

$$\mathbb{Z}_{18}^* = \{1, 5, 7, 11, 13, 17\}$$

For 1:
\begin{align*}
   1^1 = 1
\end{align*}

Therefore, the order of 1 is 1 \\

For 5:
\begin{align*}
   5^1 \equiv 5 \pmod{18} \\
   5^2 \equiv 7 \pmod{18} \\
   5^3 \equiv 17 \pmod{18} \\
   5^4 \equiv 13 \pmod{18} \\
   5^5 \equiv 11 \pmod{18} \\
   5^6 \equiv 1 \pmod{18}
\end{align*}

Therefore, the order of 5 is 6 \\

For 7:
\begin{align*}
   7^1 \equiv 7 \pmod{18} \\
   7^2 \equiv 13 \pmod{18} \\
   7^3 \equiv 1 \pmod{18}
\end{align*}

Therefore, the order of 7 is 3 \\

For 11:
\begin{align*}
   11^1 \equiv 11 \pmod{18} \\
   11^2 \equiv 13 \pmod{18} \\
   11^3 \equiv 17 \pmod{18} \\
   11^4 \equiv 7 \pmod{18} \\
   11^5 \equiv 5 \pmod{18} \\
   11^6 \equiv 1 \pmod{18}
\end{align*}

Therefore, the order of 11 is 6 \\

For 13:
\begin{align*}
   13^1 \equiv 13 \pmod{18} \\
   13^2 \equiv 7 \pmod{18} \\
   13^3 \equiv 1 \pmod{18}
\end{align*}

Therefore, the order of 13 is 3 \\

For 17:
\begin{align*}
   17^1 \equiv 17 \pmod{18} \\
   17^2 \equiv 1 \pmod{18}
\end{align*}

Therefore, the order of 17 is 2 \\

\textbf{Therefore:}
\begin{align*}
    ord(1) = 1 \\
    ord(5) = 6 \\
    ord(7) = 3 \\
    ord(11) = 6 \\
    ord(13) = 3 \\
    ord(17) = 2
\end{align*}

\newpage

\section*{Exercise 2}
Let $\psi: G \to K$ be a group isomorphism. Show that
\begin{itemize}
\item[(a)] if $G$ is abelian, then so is $K$;
\item[(b)] if $H \triangleleft G$, then $\psi(H) \triangleleft K$;
\item[(c)] $\text{ord}_G(a) = \text{ord}_K(\psi(a))$ for any $a \in G$.
\end{itemize}

\textbf{Solution:} \\

(a) Let $k_1, k_2 \in K$. Since $\psi$ is an isomorphism, there exist $g_1, g_2 \in G$ such that $\psi(g_1) = k_1$ and $\psi(g_2) = k_2$.
Then:
\begin{align*}
k_1k_2 &= \psi(g_1)\psi(g_2) \\
&= \psi(g_1g_2) \text{ (since $\psi$ is a homomorphism)} \\
&= \psi(g_2g_1) \text{ (since $G$ is abelian)} \\
&= \psi(g_2)\psi(g_1) \\
&= k_2k_1
\end{align*}

Therefore, $K$ is abelian. \\

(b) Let $k \in K$. We want to show that $k\psi(H)k^{-1} \subseteq \psi(H)$. \\

Since $\psi$ is an isomorphism, there exists $g \in G$ such that $\psi(g) = k$.
Let $h' \in \psi(H)$. Then there exists $h \in H$ such that $\psi(h) = h'$. \\

Consider:
\begin{align*}
kh'k^{-1} &= \psi(g)\psi(h)\psi(g^{-1}) \\
&= \psi(ghg^{-1}) \\
&\in \psi(H) \text{ (since $H \triangleleft G$ implies $ghg^{-1} \in H$)}
\end{align*}

Therefore, $\psi(H) \triangleleft K$. \\

(c) Let $n = \text{ord}_G(a)$. Then $a^n = e_G$ and $a^m \neq e_G$ for any $m < n$.
\begin{align*}
\psi(a^n) &= \psi(e_G) = e_K \\
\psi(a^n) &= (\psi(a))^n \text{ (by homomorphism property)}
\end{align*}

Therefore, $(\psi(a))^n = e_K$, meaning that $\text{ord}_K(\psi(a)) \leq n$. \\

Suppose there exists $m < n$ such that $(\psi(a))^m = e_K$.
Then $\psi(a^m) = e_K$. Since $\psi$ is injective, this implies $a^m = e_G$, contradicting the minimality of $n$. \\

Therefore, $\text{ord}_K(\psi(a)) = \text{ord}_G(a)$.

\newpage

\section*{Exercise 3}
Let $G$ be a finite group and $f: G \to K$ be a group homomorphism. Show that $|G| = |\text{Im } f| \cdot |\text{ker } f|$. \\

\textbf{Solution:} \\

By the First Isomorphism Theorem, we know that:
   $$G/\text{ker } f \cong \text{Im } f$$

Due to the isomorphism, we have:
\begin{align*}
   |G/\text{ker } f| &= |\text{Im } f| \\
   \frac{|G|}{|\text{ker } f|} &= |\text{Im } f|
\end{align*}

Hence:
\begin{align*}
   |G| &= |\text{Im } f| \cdot |\text{ker } f|
\end{align*}

\newpage

\section*{Exercise 4}
Show that
\begin{itemize}
\item[a)] $\text{GL}_n(\mathbb{R})/\text{SL}_n(\mathbb{R}) \simeq \mathbb{R}^\times$;
\item[b)] $\mathbb{C}^\times/C_n \simeq \mathbb{C}^\times$;
\item[c)] $\mathbb{C}/\mathbb{R} \simeq \mathbb{R}$.
\end{itemize}
(Here $\mathbb{C}$ and $\mathbb{R}$ stand for the additive groups of complex and real numbers, respectively, and $C_n = \{e^{2\pi i k/n}: k = 0,1,...,n-1\}$ stands for the multiplicative group of $n$-th roots of unity.) \\

\textbf{Solution:} \\

(a) To show $\text{GL}_n(\mathbb{R})/\text{SL}_n(\mathbb{R}) \simeq \mathbb{R}^\times$: \\

The determinant map $\det: \text{GL}_n(\mathbb{R}) \to \mathbb{R}^\times$ is a surjective group homomorphism. \\

By the First Isomorphism Theorem, we have:
$\text{GL}_n(\mathbb{R})/\ker(\det) \simeq \mathbb{R}^\times$ \\

By definition, $\ker(\det) = \text{SL}_n(\mathbb{R})$.
Therefore, $\text{GL}_n(\mathbb{R})/\text{SL}_n(\mathbb{R}) \simeq \mathbb{R}^\times$ \\

(b) To show $\mathbb{C}^\times/C_n \simeq \mathbb{C}^\times$: \\

The map $f: \mathbb{C}^\times \to \mathbb{C}^\times$ defined by $f(z) = z^n$ is a surjective group homomorphism. \\

The kernel of $f$ is precisely $C_n = \{e^{2\pi i k/n}: k = 0,1,...,n-1\}$
since these are exactly the elements that map to 1 under $f$. \\

By the First Isomorphism Theorem:
$\mathbb{C}^\times/\ker(f) \simeq \text{Im}(f) = \mathbb{C}^\times$ \\

Therefore, $\mathbb{C}^\times/C_n \simeq \mathbb{C}^\times$ \\

(c) To show $\mathbb{C}/\mathbb{R} \simeq \mathbb{R}$: \\

The map $g: \mathbb{C} \to \mathbb{R}$ defined by $g(z) = \text{Im}(z)$
where $\text{Im}(z)$ denotes the imaginary part of $z$. \\

This is a surjective group homomorphism with kernel $\mathbb{R}$
(since real numbers are precisely the complex numbers with zero imaginary part, meaning that $\ker(g) = \mathbb{R}$). \\

Therefore, $\mathbb{C}/\mathbb{R} \simeq \mathbb{R}$

\newpage

\section*{Exercise 5}
For each pair of groups from the following list, determine whether they are isomorphic or not:
\[ \mathbb{Z}_{12}, \quad \mathbb{Z}_3 \oplus \mathbb{Z}_2 \oplus \mathbb{Z}_2, \quad \mathbb{Z}_2 \oplus \mathbb{Z}_3 \oplus \mathbb{Z}_4, \quad \mathbb{Z}_3 \oplus \mathbb{Z}_4, \quad C_{12}, \quad C_4 \times C_3. \]

\textbf{Solution:} \\

The order and structure of each group:

   \begin{itemize}
   \item $|\mathbb{Z}_{12}| = 12$, cyclic
   \item $|\mathbb{Z}_3 \oplus \mathbb{Z}_2 \oplus \mathbb{Z}_2| = 3 \cdot 2 \cdot 2 = 12$, not cyclic
   \item $|\mathbb{Z}_2 \oplus \mathbb{Z}_3 \oplus \mathbb{Z}_4| = 2 \cdot 3 \cdot 4 = 24$
   \item $|\mathbb{Z}_3 \oplus \mathbb{Z}_4| = 3 \cdot 4 = 12$
   \item $|C_{12}| = 12$, cyclic
   \item $|C_4 \times C_3| = 4 \cdot 3 = 12$
   \end{itemize}

$\mathbb{Z}_2 \oplus \mathbb{Z}_3 \oplus \mathbb{Z}_4$ has order 24, while all others have order 12. Therefore, this group is not isomorphic to any other group in the list. \\

For the remaining groups of order 12:
   \begin{itemize}
   \item $\mathbb{Z}_{12}$ and $C_{12}$ are both cyclic groups of order 12, therefore they are isomorphic.
   
   \item $\mathbb{Z}_3 \oplus \mathbb{Z}_2 \oplus \mathbb{Z}_2$ has elements of order at most 6 (lcm(3,2,2) = 6), so it's not isomorphic to $\mathbb{Z}_{12}$ or $C_{12}$.
   
   \item $\mathbb{Z}_3 \oplus \mathbb{Z}_4$ has an element of order 12 (by Chinese Remainder Theorem since gcd(3,4) = 1), making it isomorphic to $\mathbb{Z}_{12}$ and $C_{12}$.
   
   \item $C_4 \times C_3$ is isomorphic to $\mathbb{Z}_{12}$ (and thus to $C_{12}$ and $\mathbb{Z}_3 \oplus \mathbb{Z}_4$) by the Chinese Remainder Theorem since gcd(4,3) = 1.
   \end{itemize}

Therefore, the groups $\mathbb{Z}_{12}, C_{12}, \mathbb{Z}_3 \oplus \mathbb{Z}_4,$ and $C_4 \times C_3$ are all isomorphic to each other.
$\mathbb{Z}_3 \oplus \mathbb{Z}_2 \oplus \mathbb{Z}_2$ and $\mathbb{Z}_2 \oplus \mathbb{Z}_3 \oplus \mathbb{Z}_4$ are not isomorphic to any other group in the list.

\newpage

\section*{Exercise 6}
Let $G$ be a group such that any non-identity element has order 2. Show that $G$ is abelian. \\

\textbf{Solution:} \\

From the question, we know that for any non-identity element $x \in G$, $x^2 = e$ (where $e$ is the identity element). \\

Let $a,b \in G$ be arbitrary elements. For the element $aba$:
\begin{align*}
(aba)^2 &= e \text{ (since every non-identity element has order 2)} \\
(aba)(aba) &= e \\
ab(aa)ba &= e \\
ab(e)ba &= e \text{ (since } a^2 = e) \\
(ab)(ba) &= e
\end{align*}

Therefore, $ba = (ab)^{-1}$ \\

However, since $(ab)^2 = e$, we know that $ab = (ab)^{-1}$ \\

Combining the two results:
$ba = (ab)^{-1} = ab$ \\

Therefore, $ab = ba$ for all $a,b \in G$, which means $G$ is abelian.

\end{document}