\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{array}
\usepackage{fancyhdr}
\usepackage{amssymb}
\usepackage[shortlabels]{enumitem}

\pagestyle{fancy}
\fancyhead[L]{Banghao Chi}
\fancyhead[C]{Homework 2}
\fancyhead[R]{7th Feb}

\fancyfoot[C]{\thepage}

\renewcommand{\headrulewidth}{0.5pt}
\renewcommand{\footrulewidth}{0.5pt}

\begin{document}

\section*{Exercise 1}
Calculate the inverses for all elements in $\mathbb{Z}_{18}^*$. \\

\textbf{Solution:} \\

First, finding all elements in $\mathbb{Z}_{18}^*$ (numbers coprime to 18):

$$\mathbb{Z}_{18}^* = \{1, 5, 7, 11, 13, 17\}$$

For $a \in \mathbb{Z}_{18}^*$, we need to find $b$ such that $ab \equiv 1 \pmod{18}$ \\

1. For $a = 1$:
  $$1 \cdot 1 \equiv 1 \pmod{18} \therefore 1^{-1} = 1$$

2. For $a = 5$:
  $$5 \cdot 11 \equiv 55 \equiv 1 \pmod{18} \therefore 5^{-1} = 11$$

3. For $a = 7$:
  $$7 \cdot 13 \equiv 91 \equiv 1 \pmod{18} \therefore 7^{-1} = 13$$

4. For $a = 11$:
  $$11 \cdot 5 \equiv 55 \equiv 1 \pmod{18} \therefore 11^{-1} = 5$$

5. For $a = 13$:
  $$13 \cdot 7 \equiv 91 \equiv 1 \pmod{18} \therefore 13^{-1} = 7$$

6. For $a = 17$:
  $$17 \cdot 17 \equiv 289 \equiv 1 \pmod{18} \therefore 17^{-1} = 17$$

Therefore, the complete list of inverses for all elements in $\mathbb{Z}_{18}^*$ is:
\begin{align*}
1^{-1} &= 1 \\
5^{-1} &= 11 \\
7^{-1} &= 13 \\
11^{-1} &= 5 \\
13^{-1} &= 7 \\
17^{-1} &= 17
\end{align*}

\newpage

\section*{Exercise 2}
Calculate
\begin{itemize}
\item[(a)] $3^{100} \pmod{10}$
\item[(b)] $5^{60} \pmod{7}$
\item[(c)] $400^{60} \pmod{61}$
\item[(d)] the last digit of $17^{50}$
\end{itemize}

\textbf{Solution:}
\begin{itemize}
\item[(a)] To find $3^{100} \pmod{10}$:
    \begin{itemize}
        \item Note that $3^4 = 81 \equiv 1 \pmod{10}$
        \item Therefore, $3^{100} \equiv 3^{4 \times 25} \equiv 1^{25} \equiv 1 \pmod{10}$
    \end{itemize}

\item[(b)] To find $5^{60} \pmod{7}$:
    \begin{itemize}
        \item By Euler's theorem, $\phi(7) = 6$
        \item Therefore, $5^6 \equiv 1 \pmod{7}$
        \item Hence, $5^{60} \equiv 5^{6 \times 10} \equiv 1^{10} \equiv 1 \pmod{7}$
    \end{itemize}

\item[(c)] To find $400^{60} \pmod{61}$:
    \begin{itemize}
        \item Note that $400 \equiv 34 \pmod{61}$ (since $400 = 6 \times 61 + 34$)
        \item Therefore, $400^{60} \equiv 34^{60} \pmod{61}$
        \item By Euler's theorem, $\phi(61) = 60$
        \item $34^{60} \equiv 1 \pmod{61}$
        \item Therefore, $400^{60} \equiv 34^{60} \equiv 1 \pmod{61}$
    \end{itemize}

\item[(d)] To find the last digit of $17^{50}$:
    \begin{itemize}
        \item This is equivalent to finding $17^{50} \pmod{10}$
        \item $17 \equiv 7 \pmod{10}$
        \item Therefore, we need to find $7^{50} \pmod{10}$
        \item By Euler's theorem, $\phi(10) = 4$
        \item Therefore, $7^4 \equiv 1 \pmod{10}$
        \item Therefore, $17^{50} \equiv 7^{50} \equiv 7^{4 \times 12 + 2} \equiv 1^{12} \times 7^2 \equiv 7^2 \equiv 9 \pmod{10}$
    \end{itemize}

\end{itemize}

\newpage

\section*{Exercise 3}
Let $\varphi: G \to H$ be a group isomorphism. Show that the inverse map $\varphi^{-1}: H \to G$ is also an isomorphism. \\

\textbf{Solution:} \\

To show that $\varphi^{-1}$ is an isomorphism, we need to prove that it is bijective and is a homomorphism. \\

\textbf{1) } $\varphi^{-1}$ is a homomorphism: \\

Let $h_1, h_2 \in H$. We need to show that $\varphi^{-1}(h_1h_2) = \varphi^{-1}(h_1)\varphi^{-1}(h_2)$ \\

Since $\varphi$ is an isomorphism, there exist unique $g_1, g_2 \in G$ such that:
$\varphi(g_1) = h_1$ and $\varphi(g_2) = h_2$ \\

Therefore, $\varphi^{-1}(h_1) = g_1$ and $\varphi^{-1}(h_2) = g_2$ \\

On the other hand, since $\varphi$ is a homomorphism:
$\varphi(g_1g_2) = \varphi(g_1)\varphi(g_2) = h_1h_2$, we get:
$\varphi^{-1}(h_1h_2) = g_1g_2$ \\

Therefore:
$\varphi^{-1}(h_1h_2) = g_1g_2 = \varphi^{-1}(h_1)\varphi^{-1}(h_2)$ \\

This shows that $\varphi^{-1}$ is a homomorphism. \\

\textbf{2) } $\varphi^{-1}$ is bijective: \\

Since $\varphi$ is an isomorphism, it is bijective. By definition of inverse function:
\begin{itemize}
    \item $\varphi^{-1}$ is injective because $\varphi$ is surjective
    \item $\varphi^{-1}$ is surjective because $\varphi$ is injective
\end{itemize}

Therefore, $\varphi^{-1}$ is a bijective homomorphism, making it an isomorphism.

\newpage

\section*{Exercise 4}
Show that the multiplicative group $C_n = \{e^{2\pi i k/n} : k = 0,1,\ldots,n-1\}$ of $n$-th roots of unity is isomorphic to $\mathbb{Z}_n$. \\

\textbf{Solution:} \\

To show that $C_n \cong \mathbb{Z}_n$, we need to prove that there exists a bijective homomorphism $\phi: \mathbb{Z}_n \to C_n$. \\

Define $\phi: \mathbb{Z}_n \to C_n$ by
\[\phi(k) = e^{2\pi i k/n}\]

where $k \in \{0,1,\ldots,n-1\}$ \\

\begin{align*}
\phi(a + b \bmod n) &= e^{2\pi i(a + b \bmod n)/n} \\
&= e^{2\pi i(a + b)/n} \\
&= e^{2\pi ia/n} \cdot e^{2\pi ib/n} \\
&= \phi(a) \cdot \phi(b)
\end{align*}

Therefore, $\phi$ is a homomorphism. \\

Suppose $\phi(a) = \phi(b)$ for some $a,b \in \{0,1,\ldots,n-1\}$
Then:
\[e^{2\pi ia/n} = e^{2\pi ib/n}\]
\[e^{2\pi i(a-b)/n} = 1\]

This means $n$ divides $(a-b)$ since these are $n$-th roots of unity.
Since $a,b \in \{0,1,\ldots,n-1\}$, we must have $a = b$, meaning$\phi$ is injective. \\

For any $e^{2\pi ik/n} \in C_n$, we have $k \in \{0,1,\ldots,n-1\}$ \\

Since $\phi(k) = e^{2\pi ik/n}$, every element in $C_n$ has a preimage in $\mathbb{Z}_n$, meaning $\phi$ is surjective. \\

Since $\phi$ is a bijective homomorphism, it is an isomorphism. Therefore, $C_n \cong \mathbb{Z}_n$.
\newpage

\section*{Exercise 5}
Suppose that $\varphi: G \to H$ is a homomorphism, and let $e_G$ and $e_H$ be the identity elements of $G$ and $H$, respectively. Show that
\begin{itemize}
\item[(a)] for all $g \in G$ and $n \in \mathbb{N}$, we have $\varphi(g^n) = (\varphi(g))^n$
\item[(b)] if $g^n = e_G$, then $\varphi(g)^n = e_H$
\end{itemize}

\textbf{Solution:} \\

(a) We prove this by mathematical induction on $n$. \\

\textbf{Base case:} For $n = 1$, we have
$\varphi(g^1) = \varphi(g) = (\varphi(g))^1$ \\

\textbf{Inductive hypothesis:} Assume that for some $k \in \mathbb{N}$, 
$\varphi(g^k) = (\varphi(g))^k$ \\

\textbf{Inductive step:} Consider $n = k + 1$
\begin{align*}
\varphi(g^{k+1}) &= \varphi(g^k \cdot g) \\
&= \varphi(g^k) \cdot \varphi(g) \quad \text{(since $\varphi$ is a homomorphism)} \\
&= (\varphi(g))^k \cdot \varphi(g) \quad \text{(by inductive hypothesis)} \\
&= (\varphi(g))^{k+1}
\end{align*}

Therefore, by mathematical induction, $\varphi(g^n) = (\varphi(g))^n$ for all $n \in \mathbb{N}$. \\

(b) Let $g \in G$ such that $g^n = e_G$. Then:
\begin{align*}
\varphi(g)^n &= \varphi(g^n) \quad \text{(by part (a))} \\
&= \varphi(e_G) \\
&= e_H \quad \text{(since homomorphisms map identity to identity)}
\end{align*}

Therefore, if $g^n = e_G$, then $\varphi(g)^n = e_H$.
\newpage

\section*{Exercise 6}
Show that a group $G$ is abelian if and only if $(ab)^2 = a^2b^2$ for all $a,b \in G$. \\

\textbf{Solution:} \\

Since this is an "if and only if" statement, we will prove both directions. \\

($\Rightarrow$) If $G$ is abelian, then $(ab)^2 = a^2b^2$ for all $a,b \in G$. \\

If $G$ is abelian, then $ab = ba$ for all $a,b \in G$. \\

Therefore:
\begin{align*}
(ab)^2 &= (ab)(ab) \\
&= a(ba)b \quad \text{(by associativity)} \\
&= a(ab)b \quad \text{(since $G$ is abelian)} \\
&= (aa)(bb) \quad \text{(by associativity)} \\
&= a^2b^2
\end{align*}

($\Leftarrow$) If $(ab)^2 = a^2b^2$ for all $a,b \in G$, then $G$ is abelian, meaning $ab = ba$. \\

For any $a,b \in G$:
\begin{align*}
(ab)^2 &= a^2b^2 \\
(ab)(ab) &= (aa)(bb) \\
ab \cdot ab &= aa \cdot bb
\end{align*}

Now, multiply both sides on the left by $(ab)^{-1}$ and on the right by $(bb)^{-1}$:
\begin{align*}
(ab)^{-1}(ab \cdot ab)(bb)^{-1} &= (ab)^{-1}(aa \cdot bb)(bb)^{-1} \\
ab \cdot (bb)^{-1} &= (ab)^{-1} \cdot aa \cdot bb \cdot (bb)^{-1} \\
ab \cdot b^{-2} &= (ab)^{-1} \cdot aa \\
\end{align*}

Multiply both sides on the right by $b$:
\begin{align*}
ab \cdot b^{-2} \cdot b &= (ab)^{-1} \cdot aa \cdot b \\
ab \cdot b^{-1} &= (ab)^{-1} \cdot aa \cdot b \\
a &= (ab)^{-1} \cdot aa \cdot b \\
a &= b^{-1}a^{-1} \cdot aa \cdot b \\
a &= b^{-1}a \cdot b
\end{align*}

Multiply both sides on the left by $b$:
\begin{align*}
ba &= a \cdot b \\
ba &= ab
\end{align*}

Hence, $G$ is abelian. \\

Therefore, we have shown that $G$ is abelian if and only if $(ab)^2 = a^2b^2$ for all $a,b \in G$.

\end{document}